### 2022-07-26 paper/localpsf.tex
more paper edits. Barely under 25 pages

The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
\begin{equation}
	\label{eq:deterministic_optimization_problem}
	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
\end{equation}
The first term in the objective function is
\begin{equation*}
	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS.
\end{equation*}
We call this term the data misfit, because it measures the mismatch between the observed data and the predicted data based on a candidate parameter $\basalfriction$. Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$. The second term,
\begin{equation*}
	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{L}(\basalfriction - \basalfriction_0)\right|^2\,dx,
\end{equation*}
is a Bilaplacian regularization term. Specifically, $\mathcal{L}$ is the inverse of the solution operator for the following elliptic PDE:
\begin{align*}
-\gamma\Delta u +\delta\,u&=s\,\,\,\,\text{ in }\Omega,\\
\gamma \nabla u\cdot\normalvec+\mu\,u&=0\,\,\,\text{ on }\partial\Omega.
\end{align*}
Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
In all results shown except one, $\gamma$ is chosen so that the solution satisfies the Morozov discrepancy principle, that is,
\begin{equation*}
	\frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \frac{1}{2}\int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
\end{equation*}
The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$. The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study the effect of changing $\gamma$ on the quality of our preconditioner. The constant $\delta$ is chosen so that the correlation length, $L=\sqrt{\gamma/\delta}$, of functions drawn from the normal distribution with covariance $\mathcal{L}^{-2}$, is $1/10$th of the radius of the domain. The function $\basalfriction_0$ is the constant function with value $\basalfriction_0(x)=10.5$.
%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.

The inverse problem is solved with a inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}. The constant function $\basalfriction_0$ is used as the initial guess. The parameter value is updated iteratively according to the formula
$
	\basalfriction_{k+1} = \basalfriction_k + \alpha_k \searchdir_k,
$
where $\searchdir_k$ approximately solves one of the following Newton linear systems
\begin{equation}
	\label{eq:newton_system}
	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn}, \searchdir_k = - g_k
\end{equation}
and the step length $\alpha_k \in \mathbb{R}$ is chosen to satisfy the Armijo conditions \cite{ARMIJO}. Here $g_k$ is the gradient of $J$, $H$ is the Hessian of $J$, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$. 
%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
We use $H_\text{gn}$ for the first five iterations, and $H$ for all subsequent iterations.  The method terminates when $||g_k||/||g_0|| < 10^{-8}$. The systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$. The PCG method only requires applying $H$ (or $H_\text{gn}$) to vectors. Each application of $H$ (or $H_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.

@@

We use the framework described in this paper to generate Hessian preconditioners. We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in $H_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches. The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$. We approximate $H_\text{gn}$ rather than $H$ because $H$ often has negative values in its integral kernel when one is away from the optimal point. Our results show that our preconditioners are also good preconditioners for $H$. The H-matrix approximation of data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $H_\text{gn}$. Straightforward calculation shows that the discretized Hessian of the regularization term is given by
$
(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
$
where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix. These matrices are sparse, so we form H-matrix representations of these matrices using standard sparse H-matrix techniques. Then we invert $\bm{M}^{-1}$, and multiply these matrices together using H-matrix methods. We factor the overall Gauss-Newton Hessian approximation, denoted $\widetilde{H}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}. The approximations for 1, 5, and 25 impulse response batches are denoted by PSF (1), PSF (5), and PSF (25), respectively.

\begin{table}
	{\small
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{3pt}
			\renewcommand{\arraystretch}{1.1}
			\begin{tabular}{c| c c c | c c c | c c c}
				&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
				\hline
				Iter & 
				\#CG & \#Stokes & $||g||$ & 
				\#CG & \#Stokes & $||g||$ & 
				\#CG & \#Stokes & $||g||$ \\
				\verb~0~ &
				\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
				\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
				\verb~1~ & \verb~4~ & \verb~2.0e+7~ \\
				\verb~1~ &
				\verb~2~ & \verb~6~ & \verb~5.8e+6~ &
				\verb~2~ & \verb~6~ & \verb~7.9e+6~ &
				\verb~2~ & \verb~6~ & \verb~5.8e+6~ \\
				\verb~2~ &
				\verb~4~ & \verb~10~ & \verb~2.4e+6~ &
				\verb~5~ & \verb~12~ & \verb~3.9e+6~ &
				\verb~4~ & \verb~10~ & \verb~2.4e+6~ \\
				\verb~3~ &
				\verb~2~ & \verb~6+22~ & \verb~5.8e+5~ &
				\verb~13~ & \verb~28~ & \verb~1.6e+6~ &
				\verb~13~ & \verb~28~ & \verb~5.8e+5~ \\
				\verb~4~ &
				\verb~5~ & \verb~12~ & \verb~5.6e+4~ &
				\verb~42~ & \verb~86~ & \verb~4.8e+5~ &
				\verb~33~ & \verb~68~ & \verb~9.1e+4~ \\
				\verb~5~ &
				\verb~10~ & \verb~22~ & \verb~3.5e+3~ &
				\verb~84~ & \verb~170~ & \verb~7.7e+4~ &
				\verb~57~ & \verb~116~ & \verb~6.2e+3~ \\
				\verb~6~ &
				\verb~14~ & \verb~30~ & \verb~2.7e+1~ &
				\verb~125~ & \verb~252~ & \verb~5.0e+3~ &
				\verb~76~ & \verb~154~ & \verb~1.1e+2~ \\
				\verb~7~ &
				\verb~0~ & \verb~2~ & \verb~4.1e-2~ &
				\verb~194~ & \verb~390~ & \verb~7.9e+1~ &
				\verb~117~ & \verb~236~ & \verb~2.3e-1~ \\
				\verb~8~ & 
				\verb~---~ & \verb~---~ & \verb~---~ &
				\verb~0~ & \verb~2~ & \verb~1.6e-1~ &
				\verb~0~ & \verb~2~ & \verb~4.1e-2~ \\
				\hline
				Total & 
				\verb~38~ & \verb~114~ & \verb~---~ &
				\verb~466~ & \verb~950~ & \verb~---~ &
				\verb~303~ & \verb~624~ & \verb~---~ \\
			\end{tabular}
			\endgroup
		\end{center}
	}
	\caption{Convergence history for solving the Stokes inverse problem using the preconditioned inexact Newton-PCG method. Preconditioners shown are our method with five impulse response batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE). The columns titled \#CG show the number of conjugate gradient iterations used to solve the Newton system at $\basalfriction_k$, for the purpose of computing a search direction towards $\basalfriction_{k+1}$. The columns titled $||g||$ show the $l^2$ norm of the gradient at $\basalfriction_k$. The columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration. This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$, plus the adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system at iteration $k$. In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner. }
	\label{tab:newton_convergence_table}
\end{table}

In Table \ref{tab:newton_convergence_table} we study the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}. We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and re-use it for all subsequent Gauss-Newton and Newton iterations. No preconditioning is used in the three Gauss-Newton iterations before the PSF (5) preconditioner is built. We compare our method with the most commonly used existing preconditioners, which are no preconditioning (denoted by NONE), and preconditioning by the regularization term in the Hessian (denoted REG). In Figure \ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner. Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning. Interestingly, we observe that no preconditioning actually outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.

@@

Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\basalfriction$. We use the preconditioned conjugate gradient method to solve a linear system with the Hessian as the coefficient operator and a right hand random vector with i.i.d. entries drawn from the standard normal distribution. In Figure \ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners. PCG converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected. PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning. In Figure \ref{fig:gamma_sweep}, we perform PCG solves on linear systems of the same form, except now we vary $\gamma$. The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s. All PSF preconditioners perform the same for medium and large values of $\gamma$. For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1). As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$. Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.

@@

In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues $\lambda$ for the generalized eigenvalue problem
$
	H u = \lambda \preconditioner u.
$
Here $H$ is the Hessian evaluated at the reconstructed parameter $\basalfriction$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle. The matrix $\preconditioner$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\basalfriction$, or the regularization Hessian. The PSF preconditioners cluster the eigenvalues of the Hessian near $1$, with more batches yielding better clustering. The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues. In Figure \ref{fig:condition_number} we show the condition number of the preconditioned Hessian $\preconditioner^{-1}H$, and preconditioned Gauss-Newton Hessian, $\preconditioner^{-1}H_\text{gn}$. The condition numbers are similar for both the Hessian and the Gauss-Newton Hessian. The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers. Regularization preconditioning and no preconditioning yield larger condition numbers, with regularization preconditioning actually increasing the condition number as compared to no preconditioning.

>>>>

The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
\begin{equation}
	\label{eq:deterministic_optimization_problem}
	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
\end{equation}
The first term in the objective function is
\begin{equation*}
	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS.
\end{equation*}
We call this term the data misfit, because it measures the mismatch between the observed data and the predicted data based on a candidate parameter $\basalfriction$. Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$. The second term,
\begin{equation*}
	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{L}(\basalfriction - \basalfriction_0)\right|^2\,dx,
\end{equation*}
is a Bilaplacian regularization term. Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{L}$ is the inverse of the solution operator for the following elliptic PDE:
\begin{align*}
-\gamma\Delta u +\delta\,u&=s\,\,\,\,\text{ in }\Omega,\\
\gamma \nabla u\cdot\normalvec+\mu\,u&=0\,\,\,\text{ on }\partial\Omega.
\end{align*}
Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
\begin{equation*}
	\frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \frac{1}{2}\int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
\end{equation*}
The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$. The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner. The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{L}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain. 
%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.

We solve the inverse problem with an inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}, using $\basalfriction_0$ as the initial guess. The parameter is updated iteratively according to the formula
$
	\basalfriction_{k+1} = \basalfriction_k + \alpha_k \searchdir_k,
$
where $\searchdir_k$ approximately solves one of the following Newton linear systems
\begin{equation}
	\label{eq:newton_system}
	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn}, \searchdir_k = - g_k
\end{equation}
and the step length $\alpha_k \in \mathbb{R}$ is chosen to satisfy the Armijo conditions \cite{ARMIJO}. Here $g_k$ is the gradient of $J$, $H$ is the Hessian, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$. 
%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
We use $H_\text{gn}$ for the first five iterations, and $H$ for all subsequent iterations.  The Newton iteration terminates when $||g_k||/||g_0|| < 10^{-8}$. Systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$. PCG only requires applying $H$ (or $H_\text{gn}$) to vectors. Each application of $H$ (or $H_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.

@@

We use the framework described in this paper to generate Hessian preconditioners. We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in $H_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches. These approximations are labeled by are denoted by PSF (1), PSF (5), and PSF (25), respectively. The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$. We approximate $H_\text{gn}$ rather than $H$ because $H$ often has negative values in its integral kernel when one is away from the optimal point. Our results show that our preconditioners are also good preconditioners for $H$. The H-matrix approximation of data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $H_\text{gn}$. Straightforward calculation shows that the discretized Hessian of the regularization term is given by
$
(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
$
where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix. These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques. We factor the overall Gauss-Newton Hessian approximation, denoted $\widetilde{H}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.

\begin{table}
	{\small
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{3pt}
			\renewcommand{\arraystretch}{1.1}
			\begin{tabular}{c| c c c | c c c | c c c}
				&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
				\hline
				Iter & 
				\#CG & \#Stokes & $||g||$ & 
				\#CG & \#Stokes & $||g||$ & 
				\#CG & \#Stokes & $||g||$ \\
				\verb~0~ &
				\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
				\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
				\verb~1~ & \verb~4~ & \verb~2.0e+7~ \\
				\verb~1~ &
				\verb~2~ & \verb~6~ & \verb~5.8e+6~ &
				\verb~2~ & \verb~6~ & \verb~7.9e+6~ &
				\verb~2~ & \verb~6~ & \verb~5.8e+6~ \\
				\verb~2~ &
				\verb~4~ & \verb~10~ & \verb~2.4e+6~ &
				\verb~5~ & \verb~12~ & \verb~3.9e+6~ &
				\verb~4~ & \verb~10~ & \verb~2.4e+6~ \\
				\verb~3~ &
				\verb~2~ & \verb~6+22~ & \verb~5.8e+5~ &
				\verb~13~ & \verb~28~ & \verb~1.6e+6~ &
				\verb~13~ & \verb~28~ & \verb~5.8e+5~ \\
				\verb~4~ &
				\verb~5~ & \verb~12~ & \verb~5.6e+4~ &
				\verb~42~ & \verb~86~ & \verb~4.8e+5~ &
				\verb~33~ & \verb~68~ & \verb~9.1e+4~ \\
				\verb~5~ &
				\verb~10~ & \verb~22~ & \verb~3.5e+3~ &
				\verb~84~ & \verb~170~ & \verb~7.7e+4~ &
				\verb~57~ & \verb~116~ & \verb~6.2e+3~ \\
				\verb~6~ &
				\verb~14~ & \verb~30~ & \verb~2.7e+1~ &
				\verb~125~ & \verb~252~ & \verb~5.0e+3~ &
				\verb~76~ & \verb~154~ & \verb~1.1e+2~ \\
				\verb~7~ &
				\verb~0~ & \verb~2~ & \verb~4.1e-2~ &
				\verb~194~ & \verb~390~ & \verb~7.9e+1~ &
				\verb~117~ & \verb~236~ & \verb~2.3e-1~ \\
				\verb~8~ & 
				\verb~---~ & \verb~---~ & \verb~---~ &
				\verb~0~ & \verb~2~ & \verb~1.6e-1~ &
				\verb~0~ & \verb~2~ & \verb~4.1e-2~ \\
				\hline
				Total & 
				\verb~38~ & \verb~114~ & \verb~---~ &
				\verb~466~ & \verb~950~ & \verb~---~ &
				\verb~303~ & \verb~624~ & \verb~---~ \\
			\end{tabular}
			\endgroup
		\end{center}
	}
	\caption{Convergence history for solving the Stokes inverse problem using preconditioned inexact Newton PCG. Preconditioners shown are our method with five impulse response batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE). Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\widehat{\basalfriction}_k$. Columns titled $||g||$ show the $l^2$ norm of the gradient at $\basalfriction_k$. Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration. This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$, plus one adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system. In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner. }
	\label{tab:newton_convergence_table}
\end{table}

Table \ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}. We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and re-use it for all subsequent Gauss-Newton and Newton iterations. No preconditioning is used in the Gauss-Newton iterations before the PSF (5) preconditioner is built. We compare our method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG). In Figure \ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner. Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.}.

@@

Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\basalfriction$. We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random i.i.d. entries drawn from the standard normal distribution. In Figure \ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners. PCG converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected. PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning. In Figure \ref{fig:gamma_sweep}, we perform PCG solves on linear systems of the same form, except now we vary $\gamma$. The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s. All PSF preconditioners perform the same for medium and large values of $\gamma$. For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1). As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$. Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.

@@

In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
$
	H u = \lambda \preconditioner u.
$
Here $H$ is the Hessian evaluated at the reconstructed parameter $\basalfriction$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle. The matrix $\preconditioner$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\basalfriction$, or the regularization Hessian. The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering. The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues. In Figure \ref{fig:condition_number} we show the condition numbers of $\preconditioner^{-1}H$ and $\preconditioner^{-1}H_\text{gn}$. The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers. Regularization preconditioning and no preconditioning yield larger condition numbers. The condition numbers are similar for both $\preconditioner^{-1}H$ and $\preconditioner^{-1}H_\text{gn}$, demonstrating that the preconditioner built based on $H_\text{gn}$ is a good preconditioner for $H$.
